New modal operators for constructive mathematics

### 2023-05-02 slides-luminy2023.tex
Absolutely final version

\usepackage[normalem]{ulem}
\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{manfnt}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning,svg.path,fadings}
\hypersetup{colorlinks=true}

@@

\begin{tikzfadingfrompicture}[name=fade bottom]
  \shade[top color=transparent!0,
    bottom color=transparent!50] (0,0) rectangle (2,2);
    \end{tikzfadingfrompicture}

@@

    \item \good{Yes}, if~$A$ is countable and membership of finitely generated ideals is decidable:
    {\footnotesize
    \begin{align*}
      \mmm_0 &= \{ 0 \}, &
      \mmm_{n+1} &= \begin{cases}
        \mmm_n + (x_n), & \text{if $1 \not\in \mmm_n + (x_n)$}, {\qquad\quad\quad\!\!}\\
        \mmm_n, & \text{else.}
      \end{cases}
    \end{align*}}
    \pause

    \item \good{Yes}, if~$A$ is countable (irrespective of membership decidability):
    {\footnotesize\begin{align*}
      \mmm_0 &= \{ 0 \}, &
      \mmm_{n+1} &= \mmm_n + (\underbrace{\{ x \in A \,|\, x = x_n \wedge
        1 \not\in \mmm_n + (x_n) \}}_{\text{a certain subsingleton set}})
    \end{align*}}
    \vspace*{3.5em}
    \visible<11>{{
      \centering
      \raisebox{0pt}[0pt][0pt]{
        \hspace*{1.5em}
        \begin{tikzpicture}
          \node (inner) at (17.3mm,-20mm) {\textit{``a bad joke''}};
          \path (0,0) pic{laurel-wreath};
        \end{tikzpicture}
        \hspace*{-3em}
        \begin{tikzpicture}
          \node (inner) at (17.5mm,-18mm) {\vbox{\small\centering\textit{``non- \\informative''}}};
          \path (0,0) pic{laurel-wreath};
        \end{tikzpicture}
      }
      \par
    }}
    \pause
    \pause
    \vspace*{-3.5em}

    \item In the general case: \bad{No}, but the \emph{first-order
    consequences} of the existence of a maximal ideal \good{do hold}.
  \end{enumerate}
\end{frame}}

@@

  {\centering\footnotesize\begin{tabular}{lll}
    \toprule
    preorder~$L$ & coverings of an element~$\sigma \in L$ & filters of~$L$ \\
    \midrule
    $X^*$ & $\{ \sigma x \,|\, x \in X \}$ & maps~$\NN \to X$ \\
    $X^*$ & $\{ \sigma x \,|\, x \in X \}$,\ \
    $\{ \sigma\tau \,|\, \tau \in X^*, a \in \sigma\tau \}$ for each~$a \in X$ & surjections~$\NN \twoheadrightarrow X$ \\
    f.g. ideals & --- & ideals \\
    f.g. ideals & $\{ \sigma+(a), \sigma+(b) \}$ for each~$ab \in \sigma$,\ \
    $\{\}$ if~$1 \in \sigma$ & prime ideals \\
    opens & $\mathcal{U}$ such that~$\sigma = \bigcup \mathcal{U}$ & points \\
    $\{\star\}$ & $\{ \star \,|\, \varphi \} \cup \{ \star \,|\, \neg\varphi \}$ &
    witnesses of~\textsc{lem}
    \\
    \bottomrule
  \end{tabular}\par}
  \bigskip

@@

\addtocounter{framenumber}{-1}
\renewcommand{\insertframeextra}{b}
{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.23cm}\includegraphics[width=\paperwidth]{topos-horses-lighter-lighter}\end{minipage}}
\begin{frame}{Constructive forcing (= Grothendieck toposes)}
  A forcing notion is a template for a \hil{forcing extension}~$V^\nabla$
  of the base universe~$V$:

@@

  \small
  \begin{multicols}{2}
    For every set~$X$, \emph{proximally} \\
    there is an enumeration~$\NN \twoheadrightarrow X$.
    \bigskip

>>>>

\usepackage[normalem]{ulem}
\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{manfnt}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning,svg.path}
\hypersetup{colorlinks=true}

@@

@@

    \item \good{Yes}, if~$A$ is countable and membership of finitely generated ideals is decidable:
    {\footnotesize
    \begin{align*}
      \mmm_0 &\defeq \{ 0 \}, &
      \mmm_{n+1} &\defeq \begin{cases}
        \mmm_n + (x_n), & \text{if $1 \not\in \mmm_n + (x_n)$}, {\qquad\quad\quad\!\!}\\
        \mmm_n, & \text{else.}
      \end{cases}
    \end{align*}}
    \pause

    \item \good{Yes}, if~$A$ is countable (irrespective of membership decidability):
    {\footnotesize\begin{align*}
      \mmm_0 &\defeq \{ 0 \}, &
      \mmm_{n+1} &\defeq \mmm_n + (\underbrace{\{ x \in A \,|\, x = x_n \wedge
        1 \not\in \mmm_n + (x_n) \}}_{\text{a certain subsingleton set}})
    \end{align*}}
    \vspace*{3.5em}
    \visible<11>{{
      \centering
      \raisebox{0pt}[0pt][0pt]{
        \hspace*{1.5em}
        \begin{tikzpicture}
          \node (inner) at (17.3mm,-20mm) {\textit{``a bad joke''}};
          \path (0,0) pic{laurel-wreath};
        \end{tikzpicture}
        \hspace*{-3em}
        \begin{tikzpicture}
          \node (inner) at (17.5mm,-18mm) {\vbox{\small\centering\textit{``non- \\informative''}}};
          \path (0,0) pic{laurel-wreath};
        \end{tikzpicture}
      }
      \par
    }}
    \pause
    \pause
    \vspace*{-3.5em}

    \item In the general case: \bad{No}\pause, \\ but \emph{first-order
    consequences} of the existence of a maximal ideal \good{do hold}.
  \end{enumerate}
\end{frame}}

@@

  {\centering\footnotesize\begin{tabular}{llll}
    \toprule
    & preorder~$L$ & coverings of an element~$\sigma \in L$ & filters of~$L$ \\
    \midrule
    \normalnumber{1} & $X^*$ & $\{ \sigma x \,|\, x \in X \}$ & maps~$\NN \to X$ \\
    \normalnumber{2} & $X^*$ & $\{ \sigma x \,|\, x \in X \}$,\ \ $\{ \sigma\tau \,|\, \tau \in X^*, a \in \sigma\tau \}$ for each~$a \in X$ & surjections~$\NN \twoheadrightarrow X$ \\
    \normalnumber{3} & f.g. ideals & --- & ideals \\
    \normalnumber{4} & f.g. ideals & $\{ \sigma+(a), \sigma+(b) \}$ for each~$ab \in \sigma$,\ \ $\{\}$ if~$1 \in \sigma$ & prime ideals \\
    \normalnumber{5} & opens & $\mathcal{U}$ such that~$\sigma = \bigcup \mathcal{U}$ & points \\
    \normalnumber{6} & $\{\star\}$ & $\{ \star \,|\, \varphi \} \cup \{ \star \,|\, \neg\varphi \}$ &
    witnesses of~\textsc{lem}
    \\
    \bottomrule
  \end{tabular}\par}
  \bigskip

@@

\addtocounter{framenumber}{-1}
\renewcommand{\insertframeextra}{b}
{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.23cm}\includegraphics[width=\paperwidth]{topos-horses-lighter-lighter}\end{minipage}}
\begin{frame}{Constructive forcing (= Grothendieck toposes)}
  A forcing notion is a template for a \hil{forcing extension}~$V^\nabla$
  of the base universe~$V$:
  \medskip

@@

  \small
  \begin{multicols}{2}
    For every inhabited set~$X$, \emph{proximally} \\
    there is an enumeration~$\NN \twoheadrightarrow X$.
    \bigskip
